A Parallel Run-Time Monitoring Architecture with Adjustable Overheads

### 2014-05-23 introduction.tex
Updates to intro and related work.

In this paper, we present an architecture that combines a hardware-based
dataflow tracking engine with a software-based monitor in order to reduce the
overheads of monitoring. By using the dataflow engine to track flows of
initialized metadata, the system is able to filter out events that do not
correspond to valid metadata, greatly reducing the amount of monitoring that
must be handled in software. For example, for an array
bounds check, the dataflow engine is able to filter out instruction that do not
correspond to pointer data. Thus, we are able to have the flexibility of
software-based monitors with greatly reduced overheads.

>>>>

Recent studies have shown that filtering out monitoring checks on clean
metadata can be used to reduce the overheads of monitoring \cite{fade-hpca14}. For example, for
an array bounds check, checking non-pointer accesses is unnecessary. In this
paper, we present an architecture that combines a hardware-based dataflow
tracking engine with a software-based monitor in order to filter out both
propagation and checking of clean metadata. By keeping a separate set of flags
to track clean metadata, we are able to handle the common case of clean
metadata without the need to read full metadata. By propagating these flags, we
are also able to handle propagation of clean metadata efficiently in hardware.
% By using reduce the
% overheads of monitoringthe dataflow engine
% to track flows of
% initialized metadata, the system is able to filter out events that do not
% correspond to valid metadata, greatly reducing the amount of monitoring that
% must be handled in software. 
Thus, we are able to have the flexibility of software-based monitors with
greatly reduced overheads.

### 2014-05-23 related.tex
Edits on Sections 1-3.
Updates to intro and related work.

% Monitoring
Monitoring

% Filtering
Hardbound filters out non-pointer data \cite{hardbound-asplos08}. FlexiTaint filters out some monitoring
and performs it in hardware. FADE provides a more general filtering engine. It
filters out redundant updates and clean checks by reading metadata values. Our
dataflow engine does not require reading the actual metadata values which may
be large. We do not handle filtering redundant updates, but we do filter out
clean checks as well as clean (i.e., uninitialized) propagations. We are able
to do this because the dataflow engine can also propagate the uninitialized
information. In addition, our dataflow engine can be used to provide adjustable
overheads without false positives which is not a capability described by FADE.

% Adjustable overheads
Adjustable overhead/cooperative debugging: Testudo \cite{testudo-micro08}, QVM
\cite{qvm-oopsla08}, Greathouse et al. \cite{greathouse-cgo11}, Huang et al.
\cite{huang-sttt12}, Lo et al. \cite{lo-rtas14}

>>>>

% Monitoring
Many monitoring schemes have been developed for various security, reliability,
debugging, and other capabilities.  As we described in
Section~\ref{sec:extensions}, there exist monitoring schemes for uninitialized
memory check \cite{mondrian-asplos02}, array bounds check
\cite{hardbound-asplos08, clause-ase07}, and dynamic information flow
tracking \cite{dift-asplos04, raksha-isca07, loki-osdi08}.
Memtracker \cite{memtracker-hpca07} uses monitoring to check for various
memory-related bugs and errors.  Control flow verification techniques perform
monitoring to check that program execution follows expected control flows
\cite{schuette-comp87, impres-dac06,
kayaalp-isca12}.  Our monitoring architecture is designed to
be generally applicable to any of these instruction-grained monitoring schemes.
In particular, our proposed architecture may be especially useful for some of
these more complex schemes which also have higher overheads.  Similarly,
although many of these schemes showed low overheads in custom hardware, our
architecture can be used to reduce the overheads when applying these schemes on
a multi-core platform.

% Filtering
Hardbound filters out non-pointer data \cite{hardbound-asplos08}. FlexiTaint filters out some monitoring
and performs it in hardware. FADE \cite{fade-hpca14} provides a more general filtering engine. It
filters out redundant updates and clean checks by reading metadata values. Our
dataflow engine does not require reading the actual metadata values which may
be large. We do not handle filtering redundant updates, but we do filter out
clean checks as well as clean (i.e., uninitialized) propagations. We are able
to do this because the dataflow engine can also propagate the uninitialized
information. In addition, our dataflow engine can be used to provide adjustable
overheads without false positives which is not a capability described by FADE.

% Sampling
Statistical sampling is a method that has been applied to various debugging
techniques in order to reduce their performance overheads. For example, Liblit
et al. sampled program runs of thousands of users in order to find bugs
\cite{liblit-pldi05}. This idea has also been applied to detecting memory
leaks~\cite{chilimbi-asplos04}, dataflow analysis~\cite{greathouse-cgo11}, and
other run-time monitoring techniques. The Testudo project
\cite{testudo-micro08} specifically targets hardware-based run-time monitors.
These sampling-based techniques all limit the debugging capabilities in order
to provide lower performance overheads, similar to our architecture's goals.
However, these approaches are not able to explicitly set an overhead budget, as
our system is able to do. In addition, with the exception of Testudo, these
methods have not considered parallel monitoring. Finally, we note that the
concept of distributing the monitoring over many users and runs to increase
coverage can also be enabled using our architecture.

% Limited Monitoring
In terms of controlling overheads, the Quality Virtual Machine (QVM) is a
modification of the Java Virtual Machine (JVM) that supports run-time
monitoring with controllable overheads \cite{qvm-oopsla08}. Similarly, Huang et
al. created a framework for controlling the overheads of software-based
monitoring \cite{huang-sttt12}. These projects are targeting a similar problem
to the one we address in this paper. However, both projects are focused on
software-based monitoring and thus their mechanisms modify the software
monitoring framework in order to enable or disable monitoring. Instead, our
architecture provides a general hardware mechanism to control overheads of
parallel monitoring architectures. In addition, these works prevented
false positives by enabling and disabling monitoring using a source-dropping
policy. Lo et al. \cite{lo-rtas14} developed a hardware architecture to limit
monitoring in hard real-time systems. Their architecture performs last-minute
dropping in order to guarantee strict deadlines. Our system allows for a
trade-off between choosing source-dropping, last-minute dropping, or min-point
dropping depending on the needs of the designer or user which has not been
explored by these previous works.
